\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R

### 2014-01-09 papers/jss/article.tex
Check in the new JSS article, and rewrite/improve my MapReduce example application section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@729 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@

%% DE: I tend to have wider option(width=...) so this
%%     guarantees better line breaks
<<echo=FALSE,print=FALSE>>=
options(width=65, prompt="R> ", digits=4)
@

@@

%% TODO(de) Can we make this not break the width of the page?
\noindent
\begin{table}
\begin{tabular}{@{\hskip .01\textwidth}p{.40\textwidth}@{\hskip .02\textwidth}@{\hskip .02\textwidth}p{0.55\textwidth}@{\hskip .01\textwidth}}
\toprule
Schema : \texttt{addressbook.proto} & Example R Session\\
\cmidrule{1-2}
\begin{minipage}{.35\textwidth}
\vspace{2mm}
\begin{example}
package tutorial;
message Person {
 required string name = 1;
 required int32 id = 2;
 optional string email = 3;
 enum PhoneType {
   MOBILE = 0; HOME = 1;
   WORK = 2;
 }
 message PhoneNumber {
   required string number = 1;
   optional PhoneType type = 2;
 }
 repeated PhoneNumber phone = 4;
}
\end{example}
\vspace{2mm}
\end{minipage} & \begin{minipage}{.5\textwidth}
<<echo=TRUE>>=
library(RProtoBuf)
p <- new(tutorial.Person, id=1, name="Dirk")
class(p)
p$name
p$name <- "Murray"
cat(as.character(p))
serialize(p, NULL)
@
\end{minipage} \\
\bottomrule
\end{tabular}
\caption{The schema representation from a \texttt{.proto} file for the
  \texttt{tutorial.Person} class (left) and simple R code for creating
  an object of this class and accessing its fields (right).}
\label{tab:proto}
\end{table}

@@

<<>>=
ls("RProtoBuf:DescriptorPool")
@

@@

<<>>=
p1 <- new(tutorial.Person)
p <- new(tutorial.Person, name = "Murray", id = 1)
@

@@

<<>>=
p$name
p$id
p$email <- "murray@stokely.org"
@

@@

<<>>=
p[["name"]] <- "Murray Stokely"
p[[ 2 ]] <- 3
p[[ "email" ]]
@

@@

<<>>=
p
@

@@

<<>>=
writeLines(as.character(p))
@

@@

<<>>=
serialize(p, NULL)
@

@@

<<>>=
tf1 <- tempfile()
serialize(p, tf1)
readBin(tf1, raw(0), 500)
@

@@

<<>>=
tf2 <- tempfile()
con <- file(tf2, open = "wb")
serialize(p, con)
close(con)
readBin(tf2, raw(0), 500)
@

@@

<<>>=
# serialize to a file
p$serialize(tf1)
# serialize to a binary connection
con <- file(tf2, open = "wb")
p$serialize(con)
close(con)
@

@@

<<>>=
msg <- read(tutorial.Person, tf1)
writeLines(as.character(msg))
@

@@

<<>>=
con <- file(tf2, open = "rb")
message <- read(tutorial.Person, con)
close(con)
writeLines(as.character(message))
@

@@

<<>>=
# reading the raw vector payload of the message
payload <- readBin(tf1, raw(0), 5000)
message <- read(tutorial.Person, payload)
@

@@

<<>>=
# reading from a file
message <- tutorial.Person$read(tf1)
# reading from a binary connection
con <- file(tf2, open = "rb")
message <- tutorial.Person$read(con)
close(con)
# read from the payload
message <- tutorial.Person$read(payload)
@

@@

<<>>=
# field descriptor
tutorial.Person$email

# enum descriptor
tutorial.Person$PhoneType

# nested type descriptor
tutorial.Person$PhoneNumber
# same as
tutorial.Person.PhoneNumber
@

@@

<<>>=
tutorial.Person$PhoneType
tutorial.Person$PhoneType$WORK
@

@@

<<>>=
f <- tutorial.Person$fileDescriptor()
f
f$Person
@

@@

<<>>=
tutorial.Person$PhoneType$value(1)
tutorial.Person$PhoneType$value(name="HOME")
tutorial.Person$PhoneType$value(number=1)
@

@@

<<echo=FALSE,print=FALSE>>=
    if (!exists("protobuf_unittest.TestAllTypes",
                "RProtoBuf:DescriptorPool")) {
        unittest.proto.file <- system.file("unitTests", "data",
                                           "unittest.proto",
                                           package="RProtoBuf")
        readProtoFiles(file=unittest.proto.file)
    }
@

<<>>=
a <- new(protobuf_unittest.TestAllTypes)
a$optional_bool <- TRUE
a$optional_bool <- FALSE
<<eval=F>>=
a$optional_bool <- NA
<<echo=FALSE,eval=TRUE,print=TRUE>>=
try(a$optional_bool <- NA,silent=TRUE)
@

@@

<<>>=
as.integer(2^31-1)
<<eval=F>>=
as.integer(2^31 - 1) + as.integer(1)
<<echo=FALSE,eval=TRUE,print=TRUE>>=
try(as.integer(2^31 - 1) + as.integer(1))

2^31
class(2^31)
@

@@

<<>>=
2^53 == (2^53 + 1)
@

@@

<<echo=FALSE,print=FALSE>>=
if (!exists("protobuf_unittest.TestAllTypes",
            "RProtoBuf:DescriptorPool")) {
    unittest.proto.file <- system.file("unitTests", "data",
                                       "unittest.proto",
                                       package="RProtoBuf")
    readProtoFiles(file=unittest.proto.file)
}
@

@@

<<>>=
test <- new(protobuf_unittest.TestAllTypes)
test$repeated_int64 <- c(2^53, 2^53+1)
length(unique(test$repeated_int64))
@

@@

<<>>=
test$repeated_int64 <- c("9007199254740992", "9007199254740993")
@

@@

<<>>=
options("RProtoBuf.int64AsString" = FALSE)
test$repeated_int64
length(unique(test$repeated_int64))
options("RProtoBuf.int64AsString" = TRUE)
test$repeated_int64
length(unique(test$repeated_int64))
@

<<echo=FALSE,print=FALSE>>=
options("RProtoBuf.int64AsString" = FALSE)
@

@@

<<echo=TRUE>>=
datasets <- subset(as.data.frame(data()$results), Package=="datasets")
datasets$load.name <- sub("\\s+.*$", "", datasets$Item)
n <- nrow(datasets)
@

There are \Sexpr{n} standard data sets included in R.  We use the
\texttt{can\_serialize\_pb} method to determine how many of those can
be safely converted to a serialized protocol buffer representation.

<<echo=TRUE>>=
#datasets$valid.proto <- sapply(datasets$load.name, function(x) can_serialize_pb(eval(as.name(x))))
#datasets <- subset(datasets, valid.proto==TRUE)
m <- nrow(datasets)
@

\Sexpr{m} data sets could be converted to Protocol Buffers
(\Sexpr{format(100*m/n,digits=1)}\%).  The next section illustrates how
many bytes were usued to store the data sets under four different
situations (1) normal R serialization, (2) R serialization followed by
gzip, (3) normal protocol buffer serialization, (4) protocol buffer
serialization followed by gzip.

@@

<<echo=FALSE,print=FALSE>>=
datasets$object.size <- unname(sapply(datasets$load.name, function(x) object.size(eval(as.name(x)))))

datasets$R.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize(eval(as.name(x)), NULL))))

datasets$R.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize(eval(as.name(x)), NULL))))

datasets$R.serialize.size.gz <- unname(sapply(datasets$load.name, function(x) length(memCompress(serialize(eval(as.name(x)), NULL), "gzip"))))

datasets$RProtoBuf.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize_pb(eval(as.name(x)), NULL))))

datasets$RProtoBuf.serialize.size.gz <- unname(sapply(datasets$load.name, function(x) length(memCompress(serialize_pb(eval(as.name(x)), NULL), "gzip"))))

clean.df <- data.frame(dataset=datasets$load.name,
                       object.size=datasets$object.size,
                       "serialized"=datasets$R.serialize.size,
                       "gzipped serialized"=datasets$R.serialize.size.gz,
                       "RProtoBuf"=datasets$RProtoBuf.serialize.size,
                       "gzipped RProtoBuf"=datasets$RProtoBuf.serialize.size.gz,
                       check.names=FALSE)
@

@@

<<eval=FALSE>>=
# Load packages
library(RProtoBuf)
library(httr)

# Retrieve and parse message
req <- GET ('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
output <- unserialize_pb(req$content)

# Check that no information was lost
identical(output, MASS::Animals)
@
This code suggests a method for exchanging objects between R servers, however this can 
also be done without protocol buffers. The main advantage of using an inter-operable format 
is that we can actually access R objects from within another
programming language. For example, in a very similar fashion we can retrieve the same
dataset in a Python client. To parse messages in Python, we first compile the 
\texttt{rexp.proto} descriptor into a python module using the \texttt{protoc} compiler:

@@

<<>>=
#requires httr >= 0.2.99
library(httr)
library(RProtoBuf)

args <- list(n=42, mean=100)
payload <- serialize_pb(args, NULL)

req <- POST (
  url = "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb",
  body = payload,
  add_headers (
    "Content-Type" = "application/x-protobuf"
  )
)

#This is the output of stats::rnorm(n=42, mean=100)
output <- unserialize_pb(req$content)
print(output)
@
The OpenCPU server basically performs the following steps to process the above RPC request:

<<eval=FALSE>>=
fnargs <- unserialize_pb(inputmsg)
val <- do.call(stats::rnorm, fnargs)
outputmsg <- serialize_pb(val)
@
In reality the OpenCPU provides a lot of meta functionality such as handling
of sessions, exceptions, security, and much more. OpenCPU also makes it possible to store
output of a function call on the server, instead of directly retrieving it. Thereby 
objects can be shared with other users or used as arguments in a subsequent
function call. But in its essence, the HTTP API provides a simple way to perform remote 
R function calls over HTTPS. The same request can be performed in Python as follows:

@@

TODO(mstokely): Make this better.

Many large data sets in fields such as particle physics and
information processing are stored in binned or histogram form in order
to reduce the data storage requirements
\citep{scott2009multivariate}. Protocol Buffers make a particularly
good data transport format in distributed MapReduces environments
where large numbers of computers process a large data set for analysis.

There are two common patterns for generating histograms of large data
sets with MapReduce.  In the first method, each mapper task can
generate a histogram over a subset of the data that is has been
assigned, and then the histograms of each mapper are sent to one or
more reducer tasks to merge.

In the second method, each mapper rounds a data point to a bucket
width and outputs that bucket as a key and '1' as a value.  Reducers
then sum up all of the values with the same key and output to a data store.

In both methods, the mapper tasks must choose identical
bucket boundaries even though they are analyzing disjoint parts of the
input set that may cover different ranges, or we must implement
multiple phases.

@@

Figure~\ref{fig:mr-histogram-pattern1} illustrates the second method
described above for histogram generation of large data sets with
MapReduce.

This package is designed to be helpful if some of the Map or Reduce
tasks are written in R, or if those components are written in other
languages and only the resulting output histograms need to be
manipulated in R.

@@

>>>>

\documentclass[article]{jss}
\usepackage{booktabs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
% Local helpers to make this more compatible with R Journal style.
%
\newcommand{\CRANpkg}[1]{\pkg{#1}}
\RequirePackage{fancyvrb}
\RequirePackage{alltt}
\DefineVerbatimEnvironment{example}{Verbatim}{}

%% almost as usual
\author{Dirk Eddelbuettel\\Debian and R Projects \And 
        Murray Stokely\\Google, Inc}
\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Dirk Eddelbuettel, Murray Stokely} %% comma-separated
\Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in R} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
Modern data collection and analysis pipelines often involve
a sophisticated mix of applications written in general purpose and
specialized programming languages.  Protocol Buffers are a popular
method of serializing structured data between applications---while remaining
independent of programming languages or operating system.  The
\CRANpkg{RProtoBuf} package provides a complete interface to this
library.
}
\Keywords{r, protocol buffers, serialization, cross-platform}
\Plainkeywords{r, protocol buffers, serialization, cross-platform} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Dirk Eddelbuettel\\
  \\
  Murray Stokely\\
  Google, Inc.\\
  1600 Amphitheatre Parkway\\
  Mountain View, CA 94040\\
  USA\\
  E-mail: \email{mstokely@google.com}\\
  URL: \url{http://www.stokely.org/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

@@

%% DE: I tend to have wider option(width=...) so this
%%     guarantees better line breaks

@@

%% TODO(de) Can we make this not break the width of the page?
\noindent
\begin{table}
\begin{tabular}{@{\hskip .01\textwidth}p{.40\textwidth}@{\hskip .02\textwidth}@{\hskip .02\textwidth}p{0.55\textwidth}@{\hskip .01\textwidth}}
\toprule
Schema : \texttt{addressbook.proto} & Example R Session\\
\cmidrule{1-2}
\begin{minipage}{.35\textwidth}
\vspace{2mm}
\begin{example}
package tutorial;
message Person {
 required string name = 1;
 required int32 id = 2;
 optional string email = 3;
 enum PhoneType {
   MOBILE = 0; HOME = 1;
   WORK = 2;
 }
 message PhoneNumber {
   required string number = 1;
   optional PhoneType type = 2;
 }
 repeated PhoneNumber phone = 4;
}
\end{example}
\vspace{2mm}
\end{minipage} & \begin{minipage}{.5\textwidth}
\begin{Schunk}
\begin{Sinput}
R> library(RProtoBuf)
R> p <- new(tutorial.Person, id=1, name="Dirk")
R> class(p)
\end{Sinput}
\begin{Soutput}
[1] "Message"
attr(,"package")
[1] "RProtoBuf"
\end{Soutput}
\begin{Sinput}
R> p$name
\end{Sinput}
\begin{Soutput}
[1] "Dirk"
\end{Soutput}
\begin{Sinput}
R> p$name <- "Murray"
R> cat(as.character(p))
\end{Sinput}
\begin{Soutput}
name: "Murray"
id: 1
\end{Soutput}
\begin{Sinput}
R> serialize(p, NULL)
\end{Sinput}
\begin{Soutput}
 [1] 0a 06 4d 75 72 72 61 79 10 01
\end{Soutput}
\end{Schunk}
\end{minipage} \\
\bottomrule
\end{tabular}
\caption{The schema representation from a \texttt{.proto} file for the
  \texttt{tutorial.Person} class (left) and simple R code for creating
  an object of this class and accessing its fields (right).}
\label{tab:proto}
\end{table}

@@

\begin{Schunk}
\begin{Sinput}
R> ls("RProtoBuf:DescriptorPool")
\end{Sinput}
\begin{Soutput}
 [1] "rexp.CMPLX"                  
 [2] "rexp.REXP"                   
 [3] "rexp.STRING"                 
 [4] "rprotobuf.HelloWorldRequest" 
 [5] "rprotobuf.HelloWorldResponse"
 [6] "tutorial.AddressBook"        
 [7] "tutorial.Person"             
 [8] "tutorial.Test1"              
 [9] "tutorial.Test2"              
[10] "tutorial.Test3"              
[11] "tutorial.Test4"              
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> p1 <- new(tutorial.Person)
R> p <- new(tutorial.Person, name = "Murray", id = 1)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> p$name
\end{Sinput}
\begin{Soutput}
[1] "Murray"
\end{Soutput}
\begin{Sinput}
R> p$id
\end{Sinput}
\begin{Soutput}
[1] 1
\end{Soutput}
\begin{Sinput}
R> p$email <- "murray@stokely.org"
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> p[["name"]] <- "Murray Stokely"
R> p[[ 2 ]] <- 3
R> p[[ "email" ]]
\end{Sinput}
\begin{Soutput}
[1] "murray@stokely.org"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> p
\end{Sinput}
\begin{Soutput}
[1] "message of type 'tutorial.Person' with 3 fields set"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> writeLines(as.character(p))
\end{Sinput}
\begin{Soutput}
name: "Murray Stokely"
id: 3
email: "murray@stokely.org"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> serialize(p, NULL)
\end{Sinput}
\begin{Soutput}
 [1] 0a 0e 4d 75 72 72 61 79 20 53 74 6f 6b 65 6c 79 10 03 1a 12
[21] 6d 75 72 72 61 79 40 73 74 6f 6b 65 6c 79 2e 6f 72 67
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> tf1 <- tempfile()
R> serialize(p, tf1)
R> readBin(tf1, raw(0), 500)
\end{Sinput}
\begin{Soutput}
 [1] 0a 0e 4d 75 72 72 61 79 20 53 74 6f 6b 65 6c 79 10 03 1a 12
[21] 6d 75 72 72 61 79 40 73 74 6f 6b 65 6c 79 2e 6f 72 67
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> tf2 <- tempfile()
R> con <- file(tf2, open = "wb")
R> serialize(p, con)
R> close(con)
R> readBin(tf2, raw(0), 500)
\end{Sinput}
\begin{Soutput}
 [1] 0a 0e 4d 75 72 72 61 79 20 53 74 6f 6b 65 6c 79 10 03 1a 12
[21] 6d 75 72 72 61 79 40 73 74 6f 6b 65 6c 79 2e 6f 72 67
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> # serialize to a file
R> p$serialize(tf1)
R> # serialize to a binary connection
R> con <- file(tf2, open = "wb")
R> p$serialize(con)
R> close(con)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> msg <- read(tutorial.Person, tf1)
R> writeLines(as.character(msg))
\end{Sinput}
\begin{Soutput}
name: "Murray Stokely"
id: 3
email: "murray@stokely.org"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> con <- file(tf2, open = "rb")
R> message <- read(tutorial.Person, con)
R> close(con)
R> writeLines(as.character(message))
\end{Sinput}
\begin{Soutput}
name: "Murray Stokely"
id: 3
email: "murray@stokely.org"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> # reading the raw vector payload of the message
R> payload <- readBin(tf1, raw(0), 5000)
R> message <- read(tutorial.Person, payload)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> # reading from a file
R> message <- tutorial.Person$read(tf1)
R> # reading from a binary connection
R> con <- file(tf2, open = "rb")
R> message <- tutorial.Person$read(con)
R> close(con)
R> # read from the payload
R> message <- tutorial.Person$read(payload)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> # field descriptor
R> tutorial.Person$email
\end{Sinput}
\begin{Soutput}
[1] "descriptor for field 'email' of type 'tutorial.Person' "
\end{Soutput}
\begin{Sinput}
R> # enum descriptor
R> tutorial.Person$PhoneType
\end{Sinput}
\begin{Soutput}
[1] "descriptor for enum 'PhoneType' of type 'tutorial.Person' with 3 values"
\end{Soutput}
\begin{Sinput}
R> # nested type descriptor
R> tutorial.Person$PhoneNumber
\end{Sinput}
\begin{Soutput}
[1] "descriptor for type 'tutorial.Person.PhoneNumber' "
\end{Soutput}
\begin{Sinput}
R> # same as
R> tutorial.Person.PhoneNumber
\end{Sinput}
\begin{Soutput}
[1] "descriptor for type 'tutorial.Person.PhoneNumber' "
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> tutorial.Person$PhoneType
\end{Sinput}
\begin{Soutput}
[1] "descriptor for enum 'PhoneType' of type 'tutorial.Person' with 3 values"
\end{Soutput}
\begin{Sinput}
R> tutorial.Person$PhoneType$WORK
\end{Sinput}
\begin{Soutput}
[1] 2
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> f <- tutorial.Person$fileDescriptor()
R> f
\end{Sinput}
\begin{Soutput}
[1] "file descriptor for package tutorial (/usr/local/google/R/RProtoBuf/proto/addressbook.proto)"
\end{Soutput}
\begin{Sinput}
R> f$Person
\end{Sinput}
\begin{Soutput}
[1] "descriptor for type 'tutorial.Person' "
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> tutorial.Person$PhoneType$value(1)
\end{Sinput}
\begin{Soutput}
[1] "enum value descriptor tutorial.Person.MOBILE"
\end{Soutput}
\begin{Sinput}
R> tutorial.Person$PhoneType$value(name="HOME")
\end{Sinput}
\begin{Soutput}
[1] "enum value descriptor tutorial.Person.HOME"
\end{Soutput}
\begin{Sinput}
R> tutorial.Person$PhoneType$value(number=1)
\end{Sinput}
\begin{Soutput}
[1] "enum value descriptor tutorial.Person.HOME"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> a <- new(protobuf_unittest.TestAllTypes)
R> a$optional_bool <- TRUE
R> a$optional_bool <- FALSE
\end{Sinput}
\end{Schunk}
\begin{Schunk}
\begin{Sinput}
R> a$optional_bool <- NA
\end{Sinput}
\end{Schunk}
\begin{Schunk}
\begin{Soutput}
[1] "Error in eval(expr, envir, enclos) : \n  NA boolean values can not be stored in bool protocol buffer fields\n"
attr(,"class")
[1] "try-error"
attr(,"condition")
<Rcpp::exception in eval(expr, envir, enclos): NA boolean values can not be stored in bool protocol buffer fields>
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> as.integer(2^31-1)
\end{Sinput}
\begin{Soutput}
[1] 2147483647
\end{Soutput}
\end{Schunk}
\begin{Schunk}
\begin{Sinput}
R> as.integer(2^31 - 1) + as.integer(1)
\end{Sinput}
\end{Schunk}
\begin{Schunk}
\begin{Soutput}
[1] NA
\end{Soutput}
\begin{Soutput}
[1] 2.147e+09
\end{Soutput}
\begin{Soutput}
[1] "numeric"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> 2^53 == (2^53 + 1)
\end{Sinput}
\begin{Soutput}
[1] TRUE
\end{Soutput}
\end{Schunk}

@@

@@

\begin{Schunk}
\begin{Sinput}
R> test <- new(protobuf_unittest.TestAllTypes)
R> test$repeated_int64 <- c(2^53, 2^53+1)
R> length(unique(test$repeated_int64))
\end{Sinput}
\begin{Soutput}
[1] 1
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> test$repeated_int64 <- c("9007199254740992", "9007199254740993")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> options("RProtoBuf.int64AsString" = FALSE)
R> test$repeated_int64
\end{Sinput}
\begin{Soutput}
[1] 9.007e+15 9.007e+15
\end{Soutput}
\begin{Sinput}
R> length(unique(test$repeated_int64))
\end{Sinput}
\begin{Soutput}
[1] 1
\end{Soutput}
\begin{Sinput}
R> options("RProtoBuf.int64AsString" = TRUE)
R> test$repeated_int64
\end{Sinput}
\begin{Soutput}
[1] "9007199254740992" "9007199254740993"
\end{Soutput}
\begin{Sinput}
R> length(unique(test$repeated_int64))
\end{Sinput}
\begin{Soutput}
[1] 2
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> datasets <- subset(as.data.frame(data()$results), Package=="datasets")
R> datasets$load.name <- sub("\\s+.*$", "", datasets$Item)
R> n <- nrow(datasets)
\end{Sinput}
\end{Schunk}

There are 103 standard data sets included in R.  We use the
\texttt{can\_serialize\_pb} method to determine how many of those can
be safely converted to a serialized protocol buffer representation.

\begin{Schunk}
\begin{Sinput}
R> #datasets$valid.proto <- sapply(datasets$load.name, function(x) can_serialize_pb(eval(as.name(x))))
R> #datasets <- subset(datasets, valid.proto==TRUE)
R> m <- nrow(datasets)
\end{Sinput}
\end{Schunk}

103 data sets could be converted to Protocol Buffers
(100\%).  The next section illustrates how
many bytes were usued to store the data sets under four different
situations (1) normal R serialization, (2) R serialization followed by
gzip, (3) normal protocol buffer serialization, (4) protocol buffer
serialization followed by gzip.

@@

@@

\begin{Schunk}
\begin{Sinput}
R> # Load packages
R> library(RProtoBuf)
R> library(httr)
R> # Retrieve and parse message
R> req <- GET ('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
R> output <- unserialize_pb(req$content)
R> # Check that no information was lost
R> identical(output, MASS::Animals)
\end{Sinput}
\end{Schunk}
This code suggests a method for exchanging objects between R servers, however this can 
also be done without protocol buffers. The main advantage of using an inter-operable format 
is that we can actually access R objects from within another
programming language. For example, in a very similar fashion we can retrieve the same
dataset in a Python client. To parse messages in Python, we first compile the 
\texttt{rexp.proto} descriptor into a python module using the \texttt{protoc} compiler:

@@

\begin{Schunk}
\begin{Sinput}
R> #requires httr >= 0.2.99
R> library(httr)
R> library(RProtoBuf)
R> args <- list(n=42, mean=100)
R> payload <- serialize_pb(args, NULL)
R> req <- POST (
+   url = "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb",
+   body = payload,
+   add_headers (
+     "Content-Type" = "application/x-protobuf"
+   )
+ )
R> #This is the output of stats::rnorm(n=42, mean=100)
R> output <- unserialize_pb(req$content)
R> print(output)
\end{Sinput}
\begin{Soutput}
 [1] 100.02 100.75  99.91  98.92  97.19  98.07 102.34  98.48
 [9] 100.31 100.58 100.65 100.64  98.02  97.91 100.57 100.41
[17] 100.11 103.32 100.88 100.48 101.06 100.52 101.58 101.23
[25] 100.73 100.40  98.49 101.50  99.32  99.12  99.57  98.80
[33]  98.60  99.83  99.62 100.43 100.24 100.30  99.13  99.56
[41]  99.78 101.01
\end{Soutput}
\end{Schunk}
The OpenCPU server basically performs the following steps to process the above RPC request:

\begin{Schunk}
\begin{Sinput}
R> fnargs <- unserialize_pb(inputmsg)
R> val <- do.call(stats::rnorm, fnargs)
R> outputmsg <- serialize_pb(val)
\end{Sinput}
\end{Schunk}
In reality the OpenCPU provides a lot of meta functionality such as handling
of sessions, exceptions, security, and much more. OpenCPU also makes it possible to store
output of a function call on the server, instead of directly retrieving it. Thereby 
objects can be shared with other users or used as arguments in a subsequent
function call. But in its essence, the HTTP API provides a simple way to perform remote 
R function calls over HTTPS. The same request can be performed in Python as follows:

@@

The MapReduce programming model \citep{dean2008mapreduce} has emerged
in the last decade as a popular framework for processing large data
sets in parallel on large compute clusters.  Protocol Buffers
provide a convenient mechanism to send structured data between tasks
in a MapReduce cluster.  In particular, the large data sets in fields
such as particle physics and information processing are frequently
stored in binned or histogram form in order to reduce the data storage
requirements for later data analysis \citep{scott2009multivariate}.

In such environments, analysts may be interested in very rare
phenomenon or be dealing with highly skewed data sets or inflexible
raw data storage systems from which unbiased sampling is not feasible.
There are two common patterns for generating histograms of large data
sets in a single pass with MapReduce.  In the first method, each
mapper task generates a histogram over a subset of the data that it
has been assigned, serializes this histogram and sends it to one or
more reducer tasks which merge the intermediate histograms from the
mappers.

In the second method, illustrated in
Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data
point to a bucket width and outputs that bucket as a key and '1' as a
value.  Reducers then sum up all of the values with the same key and
output to a data store.

In both methods, the mapper tasks must choose identical bucket
boundaries in advance if we are to construct the histogram in a single
pass, even though they are analyzing disjoint parts of the input set
that may cover different ranges.  The \pkg{HistogramTools} package
\citep{histogramtools} enhances \pkg{RProtoBuf} by providing a concise
schema for R histogram objects.  The histogram message type is
designed to be helpful if some of the Map or Reduce tasks are written
in R, or if those components are written in other languages and only
the resulting output histograms need to be manipulated in R.

@@

@@

%\section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\end{document}
